Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis

### 2015-07-15 ebayes_rnaseq_heterosis.tex
typos

We used the methods described at the beginning of Section \ref{s:ebayes} to obtain normalization factors $\hat{\gamma}$ and gene-specific parameter estimates $\hat{\theta}_g$ for all genes using the \edgeR{} package \citep{robinson2010edgeR} from Bioconductor \citep{gentleman2004bioconductor} applied to the maize data on 27,888 genes with average count at least one and at most two zero read counts for each variety across the four replicates.  This analysis produced sample-specific normalization factors of $\hat{\gamma}=
(0.074, -0.059, -0.074, -0.014, -0.014, -0.124,  0.093,  0.063,  0.021, -0.037,  0.049,  0.021)$. The  gene-specific parameter estimates were treated as the true parameter values for the simulation study so that our simulated datasets mimicked the existing structure among the gene-variety means of the original maize data. If $\hat{\delta}_g<-|\hat{\alpha}_g$ ($\hat{\delta}_g >|\hat{\alpha}_g|$)  the gene was considered to have LPH (HPH).

@@

Figure \ref{f:volcano} provides a volcano plot, in this case a bivariate histogram, to visualize the maximum of the probabilities of LPH and HPH versus estimated effect size. 
\begin{figure}
\centerline{\includegraphics[width=\textwidth]{volcano}}
\caption{A bivariate histogram of the maximum of the LPH and HPH probabilities versus estimated effect size defined in equation \ref{e:effect_size} for the B73 $\times$ Mo17 maize experiment.}
\label{f:volcano}
\end{figure}
The figure shows a peak at zero for probabilities below 0.5 with estimated effect sizes of zero. Above a probability of 0.5, we see a prototypical volcano pattern, with increased heterosis probability corresponding to larger estimated effect sizes and no estimated effect sizes near zero for genes with high heterosis probability. We also see asymmetry, with larger negative effect sizes than positive effect sizes due to genes with hybrid counts near zero and relatively high parental counts. Genes with high estimated posterior probabilities of heterosis and large estimated effect sizes are candidates for further investigation.

>>>>

We used the methods described at the beginning of Section \ref{s:ebayes} to obtain normalization factors $\hat{\gamma}$ and gene-specific parameter estimates $\hat{\theta}_g$ for all genes using the \edgeR{} package \citep{robinson2010edgeR} from Bioconductor \citep{gentleman2004bioconductor} applied to the maize data on 27,888 genes with average count at least one and at most two zero read counts for each variety across the four replicates.  This analysis produced sample-specific normalization factors of $\hat{\gamma}=
(0.074, -0.059, -0.074, -0.014, -0.014, -0.124,  0.093,  0.063,  0.021, -0.037,  0.049,  0.021)$. The  gene-specific parameter estimates were treated as the true parameter values for the simulation study so that our simulated datasets mimicked the existing structure among the gene-variety means of the original maize data. If $\hat{\delta}_g<-|\hat{\alpha}_g|$ ($\hat{\delta}_g >|\hat{\alpha}_g|$)  the gene was considered to have LPH (HPH).

@@

Figure \ref{f:volcano} provides a volcano plot, in this case a bivariate histogram, to visualize the maximum of the probabilities of LPH and HPH versus estimated effect size. 
\begin{figure}
\centerline{\includegraphics[width=\textwidth]{volcano}}
\caption{A bivariate histogram of the maximum of the LPH and HPH probabilities versus estimated effect size defined in equation \eqref{e:effect_size} for the B73 $\times$ Mo17 maize experiment.}
\label{f:volcano}
\end{figure}
The figure shows a peak at zero for probabilities below 0.5 with estimated effect sizes of zero. Above a probability of 0.5, we see a prototypical volcano pattern, with increased heterosis probability corresponding to larger estimated effect sizes and no estimated effect sizes near zero for genes with high heterosis probability. We also see asymmetry, with larger negative effect sizes than positive effect sizes due to genes with hybrid counts near zero and relatively high parental counts. Genes with high estimated posterior probabilities of heterosis and large estimated effect sizes are candidates for further investigation.
